Dependent Types for Safe and Secure Web Programming

### 2014-02-03 paper/main.tex
Paper revisions

\documentclass[preprint]{sigplanconf}

@@

@@

\begin{abstract}
Dependently-typed languages allow precise types to be used during development,
facilitating static reasoning about program behaviour. However,
stronger types bring a disadvantage that it becomes increasingly difficult to
write programs that are accepted by a type checker, meaning additional proofs may
have to be specified by a developer.

Embedded domain-specific languages (EDSLs) can help address this problem by
introducing a layer of abstraction over more precise underlying types,
allowing domain-specific code to be written in a high-level language which uses
dependent types to enforce invariants without imposing additional proof
obligations on an application developer.

In this paper, we apply this technique to web programming.  Using the
dependently typed programming language \Idris{}, we introduce an EDSL to
facilitate the creation and handling of \emph{statically checked} web forms,
reducing the scope for programmer error and attacks such as SQL injection. We
also show how to enforce resource usage protocols associated with common web
operations such as CGI, database access and session handling.

@@

%\category{CR-number}{subcategory}{third-level}

@@

\keywords
Dependent types, web applications, security, verification
%keyword1, keyword2

@@

\acks
This work has been supported by the Scottish Informatics and Computer Science Alliance (SICSA) and the EPSRC. We would like to thank the contributors to the \idris{} language, especially the authors of the original \texttt{Network.Cgi} and \texttt{SQLite} libraries.
We are very grateful to Peter Thiemann and the anonymous reviewers for their insightful and constructive comments.
% SICSA / EPSRC (grant number?
% #idris
% Idris contributors, in particular Melissa for the SQLite bindings and whoever wrote Network.Cgi

>>>>

\documentclass[]{sigplanconf}

@@

\authorversion

@@

\begin{abstract}
Dependently-typed languages allow precise types to be used during development,
facilitating static reasoning about program behaviour. However,
with the use of more specific types comes the disadvantage that it becomes increasingly difficult to
write programs that are accepted by a type checker, meaning additional proofs may have to be specified manually.

Embedded domain-specific languages (EDSLs) can help address this problem by
introducing a layer of abstraction over more precise underlying types,
allowing domain-specific code to be written in a verified high-level language without imposing additional proof
obligations on an application developer.

In this paper, we apply this technique to web programming.  Using the
dependently typed programming language \Idris{}, we show how to use EDSLs to enforce resource usage protocols associated with common web operations such as CGI, database access and session handling. We also introduce an EDSL which uses dependent types to facilitate the creation and handling of web forms, reducing the scope for programmer error and possible security implications.

@@

\category{D.3.2}{Programming Languages}{Language Classifications---Applicative (functional) Languages}

@@

\keywords
Dependent Types, Web Applications, Verification
%keyword1, keyword2

@@

\acks
This work has been supported by the Scottish Informatics and Computer Science Alliance (SICSA) and the EPSRC. We would like to thank the contributors to the \idris{} language, especially the authors of the original \texttt{Network.Cgi} and \texttt{SQLite} libraries.
We are very grateful to Peter Thiemann and the anonymous reviewers for their insightful and constructive comments and suggestions.
% SICSA / EPSRC (grant number?
% #idris
% Idris contributors, in particular Melissa for the SQLite bindings and whoever wrote Network.Cgi

### 2014-02-03 paper/protocols.tex

CGI is used to invoke an application on a web server, making use of environment
variables to convey information gained from an HTTP request and using standard
output to communicate with the remote client. Importantly, HTTP headers must be
correctly written to the browser prior to any other output; failure to do so
will result in an internal server error being shown.

@@

\noindent
This is then passed, along with initial values for other effects that the user
may wish to use, to the runAction function, which invokes the RunAction
operation and executes the user-specified action.
%
A simple ``Hello, world!'' program would be defined as follows:

@@

\noindent
In order to provide the necessary static guarantee that there is a valid row to
process in the buffer, we make use of the \texttt{if\_valid} construct. The
\texttt{nextRow} function will either fetch a row into the buffer, or indicate
that a row could not be loaded, as shown by its type:

>>>>

The Common Gateway Interface (CGI) is used to invoke an application on a web server, making use of environment
variables to convey information gained from an HTTP request and using standard
output to communicate with the remote client. Importantly, HTTP headers must be
correctly written to the browser prior to any other output; failure to do so
will result in an HTTP Error 500.

@@

\noindent
This is then passed, along with initial values for other effects that the user
may wish to use, to the \texttt{runAction} function, which invokes the \texttt{RunAction}
operation and executes the user-specified action.
%
A simple ``Hello, world!'' program would be defined as follows:

@@

\noindent
In order to provide the necessary static guarantee that there is a valid row to
process in the buffer, we make use of the \texttt{if\_valid} construct, as described in Section ~\ref{rp-effects}. The
\texttt{nextRow} function will either fetch a row into the buffer, or indicate
that a row could not be loaded, as shown by its type:
